Monte Carlo Synthetic Acceleration Solutions for the Simplified $P_N$ Approximation

### 2014-09-17 doc/ttsp2013/paper/ttsp_2013.tex
adding c5g7 7-group cross sections

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery, Steven P. Hamilton, Thomas M. Evans, and Paul
  P.H. Wilson
  \\ \href{mailto:slatterysr@ornl.gov}{\texttt{slatterysr@ornl.gov}} }

@@

We present Monte Carlo Synthetic Acceleration (MCSA) as a linear
solver technique for the simplified $P_N$ ($SP_N$) discretization of
the Boltzmann neutron transport equation. Using the fully-formed
linear operator for the transport problem, we explore solutions to the
$SP_N$ equations with MCSA using criticality calculations driving
problem. Due to neutron scattering in the light water moderator,
several difficulties arise when applying MCSA to these problems that
were not observed in previous work where the method was applied to
more opaque thermal radiaton transport systems. Of most significance,
it was discovered for these problems that simply reducing the spectral
radius of the Monte Carlo problem below unity is not enough to
efficiently achieve convergence as reported in previous work. In order
to effectively solve the $SP_N$ equations for the model problem, an
algebraic preconditioning strategy is developed for MCSA where we
utilize sparse approximate inverse representations of the operator as
preconditioners to achieve convergence. Using the preconditioners,
MCSA performance is compared with production implementations of
preconditioned Krylov methods. Our studies indicate that
the iterative performance of MCSA when used to solve the $SP_N$
equations for light water reactor problems is competitive with the
preconditioned Krylov methods when operator spectrum is properly
conditioned. To solve large problems, we have parallelized the method
by constructing a variant of the multiple-set overlapping-domain
(MSOD) algorithm for Monte Carlo calculations. We discuss the method
and then present results indicating its scalability.

>>>>

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery, Steven P. Hamilton, and Thomas M. Evans
  \\ \href{mailto:slatterysr@ornl.gov}{\texttt{slatterysr@ornl.gov}} }

@@

We present Monte Carlo Synthetic Acceleration (MCSA) as a linear
solver technique for the simplified $P_N$ ($SP_N$) discretization of
the Boltzmann neutron transport equation. Using the fully-formed
linear operators for the transport problem, we explore solutions to
the $SP_N$ equations with MCSA using criticality calculations as the
driving problems. Due to neutron scattering in the light water
moderator, several difficulties arise when applying MCSA to these
problems that were not observed in previous work where the method was
applied to more opaque thermal radiaton transport systems. Of most
significance, it was discovered for these problems that simply
reducing the spectral radius of the Monte Carlo problem below unity is
not enough to efficiently achieve convergence as reported in previous
work. In order to effectively solve the $SP_N$ equations for the model
problem, an algebraic preconditioning strategy is developed for MCSA
where we utilize sparse approximate inverse representations of the
operator as preconditioners to achieve convergence. Using the
preconditioners, MCSA performance is compared with production
implementations of preconditioned Krylov methods. Our studies indicate
that the iterative performance of MCSA when used to solve the $SP_N$
equations for light water reactor problems is competitive with the
preconditioned Krylov methods when operator spectrum is properly
conditioned. To solve large problems, we have parallelized the method
by constructing a variant of the multiple-set overlapping-domain
(MSOD) algorithm for Monte Carlo calculations. We discuss the method
and then present results indicating its scalability.
